Happiness versus Income Awareness

### 2022-02-28 outputs/paper/Paper2-copy.Rmd
step3

\newpage
# Introduction
After the introduction of easy internet access to Norwegian tax data in 2001, it became possible for everyone in the nation to quickly and easily see the earnings of anyone else in the country. In Norway, everyone was just a few clicks away from discovering the average income of everyone else in the nation. In the busiest week of the year, these websites were more popular than YouTube, according to data.Because income transparency makes it easier to compare incomes, it has the potential to expand the disparity in well-being between wealthier and poorer people. Individuals with lower incomes often lose in this game of income comparisons.

Based on information gathered by the Norwegian Monitor Survey, an annual self-completion questionnaire was disseminated to an ethnically diverse sample of Norwegians between 1985 and 2013. In terms of several apparent qualities, this group seems to be typical of the broader population. According to one website, there were 29.4 million searches made in the first year after the 2007 tax data (VG 2008). The Norwegian Ministry of Finance (2014) reported 920,896 unique users completed approximately more than 17 million searches in the previous year.A total of 3,797,822 persons, or more than 24.25 percent of Norway's adult population, looked for at least one tax record in 2013, with the average user doing 18.46 searches. The mismatch may be attributed to the 2011 requirement that users check in to the official tax office website in order to search for tax data.

@@

## Literature Review
Many researches have studies about the relationship between income transparency and happiness. Many suggested one's feeling of well-being is influenced by one's relative wealth. These findings contribute to the understanding of the Easterly Paradox, which states that an increase in wealth for everyone does not result in an increase in general happiness (PNAS 2010). Because utility is a matter of perspective, policymakers have been pushed to consider factors other than GDP. The so-called resource curse,' which refers to the propensity of resource-rich nations to underperform in a variety of social-economic outcomes, has piqued the public's attention in recent years. Recent research has gone beyond the usual influence on economic growth and instead examined the implications on wider indices of human well-being, such as infant mortality. In recent decades, a distinct empirical literature has investigated the factors of happiness and subjective well-being in more depth (using either country or household data).

In the paper "Happiness, Income, and Poverty," the author first presents some evidence that well-being is unquestionably related to income, but then explores two factors that imply GDP per capita may still play a role in happiness-based policy: subjective well-being inequality and the specific example of individuals who are in income poverty, both of which imply GDP per capita may still play a role in happiness-based policy (Clark 2017). Income inequality is a new component that is introduced in the study "The Effects of Income Levels and Income Inequalities on Happiness" by Tavor et al. (2017), which examines the relationship between happiness and income inequality. Happiness has been shown to have a negative impact at severe levels of inequality defined by the Gini index, independent of GDPppp per capita. This has been experimentally shown. On the other hand, across intermediate ranges of the Gini index, the relationship between index changes and pleasure is equivocal, as seen in Figure 1. These conclusions may be drawn independently of the actual figures of GDPppp per capita in the country. Bringing these two empirical lines of study together results in the publication "Happiness and the Resource Curse," by Ali et al. (2019), which is published in the Journal of Happiness Studies. This is the first research that we are aware of that takes use of a large panel dataset to investigate the relationships between changes in happiness across nations and different metrics of resource richness, according to the authors. Our findings, which are consistent with past empirical evidence of a resource curse in oil-rich countries, show that increases in oil rents are adversely related to increases in happiness over time. This 'resource curse' curse on happiness seems to be oil-specific, and it appears to hold true for both the levels and changes in levels of happiness.

@@

Subject Well-Being -- the main outcome of interest. The Norwegian Monitor Survey contains questions about happiness and life satisfaction, which are the two most extensively used measures of subjective well-being. (Easterlin 2004, Kahneman and Deaton 2010).

@@

Satisfaction: The life satisfaction question is, “How satisfied are you with your life? Very satisfied; Somewhat Satisfied; Neither satisfied nor dissatisfied; Slightly dissatisfied; Very dissatisfied.” Researchers of the original paper "The effects of income transparency on well-being" (Ricardo Perez-Truglia, 2020) employ the Probit-OLS method to assign these values (van Praag and Ferrer-i-Carbonell 2008).

@@

@@

Norwegian had access to tax information, as shown by documentation included in the original reproduction package. Perhaps the most compelling evidence comes from a Synovate poll done in 2007, which was representative of the whole taxable population of the United States at the time. Statistics on web traffic show that online tax listings are quite popular, which supports media reporting on their tremendous popularity in this area. According to one website, there were 29.4 million searches made in the first year after the publication of the 2007 tax data (which was published in January) (VG 2008). On top of all of that, there is information that is publicly available from a time when tax records could only be accessed via the tax agency's official website. As reported by the Norwegian Ministry of Finance (2014), 920,896 unique users completed slightly more than 17 million questions in the previous year, according to the most recent available data. 16 During that specific calendar year, only adults with a valid account were able to do searches on the official website. A total of 3,797,822 persons, or more than 24.25 percent of Norway's adult population, looked for at least one tax record in 2013, with the average user doing 18.46 searches. It is not feasible to directly compare figures from 2007 and 2013 since they are derived from a range of sources and are most likely based on different definitions, making direct comparison impossible. As a consequence of this restriction, the number of people doing searches, as well as the number of searches per capital, fell between 2007 and 2013. In most cases, this mismatch may be attributed to the 2011 requirement that users check in to the official tax office website in order to search for tax data, which was put in place in 2011.

@@

To measure the popularity of the tax lists, we rely on three sources of information. The most concrete evidence comes from a Synovate poll performed in 2007, which was representative of the taxable population. The web traffic data backs up media statements about the tremendous popularity of online tax lists. One website claims 29.4 million searches in the year after the publication of the 2007 tax records (VG 2008). There are also publicly available statistics from the time when tax records could only be accessed via the tax agency's official website. According to the Norwegian Ministry of Finance (2014), in 2013, 920,896 unique users completed slightly more than 17 million queries. 16 Only adults with a valid account were able to do searches on the official website that year. In 2013, over 24.25 percent of the 3,797,822 adults in Norway looked for at least one tax record, with the average user doing 18.46 searches. Statistics from 2007 and 2013 are not directly comparable since they originate from various sources and are most likely based on different definitions. With that proviso in mind, the number of people undertaking searches and the number of searches per capital declined between 2007 and 2013. This disparity is most likely related to the 2011 requirement that users check in to the official tax office website in order to look for tax data.

@@

Table 1 shows the result of regression model using the same OLS regression. Given the dependent variable happiness, we can see that the estimated coefficient on Income Rank is 0.120, which is positive. Moreover, the p-value of the coefficient is smaller than 0.05, which indicates that Income Rank as explanatory variable is statistical significant to our dependent variable(happiness). This coefficient implies that going from the lowest to the highest income rank in Norway during 1985–2000 was associated with an increase in happiness of 0.120 standard deviations. In addition, this table reports that the coefficient of the interaction term is 0.090. The p-value of 0.032 also implies that this variable is statistical significant to our dependent variable happiness.

@@

@@

Card, David, Alexandre Mas, Enrico Moretti, and Emmanuel Saez. 2012. “Inequality at Work: The
Effect of Peer Salaries on Job Satisfaction.” American Economic Review 102 (6): 2981–3003.

NRK. 2008. “Dette er økonomisk pornografi!,” October 14.

Norway Ministry of Finance. 2014. “Høringsnotat om innstramminger i reglene om offentlige skat-
telister.” January 24, 2014.

Andersen, Ingunn, and Stian Eisenträger. 2008. “I natt kommer de dye skattetallene: 29 millioner søk I skattelistene på VG Nett,” VG, October 14.

Norway Ministry of Finance. 2014. “Høringsnotat om innstramminger i reglene om offentlige skat-
telister.” January 24, 2014.

NRK. 2008. “Dette er økonomisk pornografi!,” October 14.

Omdahl, Jan. 2009. “Nå får du skattesøk på iPhone,” Dagbladet, October 15.

Origo. 2010. “Debatten om skatten,” October 20.

Clark, A.E. Happiness, income and poverty. Int Rev Econ 64, 145–158 (2017). https://doi.org/10.1007/s12232-017-0274-7,

Tavor, T., Gonen, L. D., Weber, M., &amp; Spiegel, U. (2017). The effects of income levels and income inequalities on happiness. Journal of Happiness Studies, 19(7), 2115–2137. https://doi.org/10.1007/s10902-017-9911-9

Ali, S., Murshed, S.M. & Papyrakis, E. Happiness and the Resource Curse. J Happiness Stud 21, 437–464 (2020). https://doi.org/10.1007/s10902-019-00080-3

PNAS December 28, 2010 107 (52) 22463-22468; https://doi.org/10.1073/pnas.1015962107
Contributed by Richard A. Easterlin, October 26, 2010 (sent for review October 1, 2010)

>>>>

\newpage
# Introduction

After the introduction of easy internet access to Norwegian tax data in 2001, it became possible for everyone in the nation to quickly and easily see the earnings of anyone else in the country. In Norway, everyone was just a few clicks away from discovering the average income of everyone else in the nation. In the busiest week of the year, these websites were more popular than YouTube. Because income transparency makes it easier to compare incomes, it has the potential to expand the disparity in well-being between wealthier and poorer people. Individuals with lower incomes often lose in this game of income comparisons.

Based on information gathered by the Norwegian Monitor Survey, an annual self-completion questionnaire was disseminated to an ethnically diverse sample of Norwegians between 1985 and 2013. In terms of several apparent qualities, this group seems to be typical of the broader population. According to one website, there were 29.4 million searches made in the first year after the 2007 tax data [@citeAnderson]. @citenorway reported 920,896 unique users completed approximately more than 17 million searches. The mismatch may be attributed to the 2011 requirement that users check in to the official tax office website in order to search for tax data.

@@

## Literature Review
Many researches have studies about the relationship between income transparency and happiness. Many suggested one's feeling of well-being is influenced by one's relative wealth. These findings contribute to the understanding of the Easterly Paradox, which states that an increase in wealth for everyone does not result in an increase in general happiness[@citePNAS]. Because utility is a matter of perspective, policymakers have been pushed to consider factors other than GDP. The so-called resource curse,' which refers to the propensity of resource-rich nations to underperform in a variety of social-economic outcomes, has piqued the public's attention in recent years. Recent research has gone beyond the usual influence on economic growth and instead examined the implications on wider indices of human well-being, such as infant mortality. In recent decades, a distinct empirical literature has investigated the factors of happiness and subjective well-being in more depth (using either country or household data).

In the paper "Happiness, Income, and Poverty," the author first presents some evidence that well-being is unquestionably related to income, but then explores two factors that imply GDP per capita may still play a role in happiness-based policy: subjective well-being inequality and the specific example of individuals who are in income poverty, both of which imply GDP per capita may still play a role in happiness-based policy[@citeClark]. Income inequality is a new component that is introduced in the study "The Effects of Income Levels and Income Inequalities on Happiness" by @citeTavor, which examines the relationship between happiness and income inequality. Happiness has been shown to have a negative impact at severe levels of inequality defined by the Gini index, independent of GDPppp per capita. This has been experimentally shown. On the other hand, across intermediate ranges of the Gini index, the relationship between index changes and pleasure is equivocal, as seen in Figure 1. These conclusions may be drawn independently of the actual figures of GDPppp per capita in the country. Bringing these two empirical lines of study together results in the publication "Happiness and the Resource Curse," by @citeAli, which is published in the Journal of Happiness Studies. This is the first research that we are aware of that takes use of a large panel dataset to investigate the relationships between changes in happiness across nations and different metrics of resource richness, according to the authors. Our findings, which are consistent with past empirical evidence of a resource curse in oil-rich countries, show that increases in oil rents are adversely related to increases in happiness over time. This 'resource curse' curse on happiness seems to be oil-specific, and it appears to hold true for both the levels and changes in levels of happiness.

@@

Subject Well-Being -- the main outcome of interest. The Norwegian Monitor Survey contains questions about happiness and life satisfaction, which are the two most extensively used measures of subjective well-being [citeEaster].

@@

Satisfaction: The life satisfaction question is, “How satisfied are you with your life? Very satisfied; Somewhat Satisfied; Neither satisfied nor dissatisfied; Slightly dissatisfied; Very dissatisfied.” Researchers of the original paper.

@@

Linear regression

Dummy variable

Control variable

Coefficient

@@

Perhaps the most compelling evidence comes from a Synovate poll done in 2007, which was representative of the whole taxable population of the United States at the time. Statistics on web traffic show that online tax listings are quite popular, which supports media reporting on their tremendous popularity in this area. According to one website, there were 29.4 million searches made in the first year after the publication of the 2007 tax data (which was published in January) [@citeAnderson]. On top of all of that, there is information that is publicly available from a time when tax records could only be accessed via the tax agency's official website. As reported by the @citenorway, 920,896 unique users completed slightly more than 17 million questions in the previous year, according to the most recent available data. 16 During that specific calendar year, only adults with a valid account were able to do searches on the official website. A total of 3,797,822 persons, or more than 24.25 percent of Norway's adult population, looked for at least one tax record in 2013, with the average user doing 18.46 searches. It is not feasible to directly compare figures from 2007 and 2013 since they are derived from a range of sources and are most likely based on different definitions, making direct comparison impossible. As a consequence of this restriction, the number of people doing searches, as well as the number of searches per capital, fell between 2007 and 2013. In most cases, this mismatch may be attributed to the 2011 requirement that users check in to the official tax office website in order to search for tax data, which was put in place in 2011.

@@

Data from Google Trends may be used to compare the popularity of the income search tool with that of other similar services. It is taken into consideration when determining the popularity of the income search tool how many times a phrase has been searched in the Google search engine. In the main search category, skattelister, we include searches for the two words that are most frequently used to refer to tax records: "skattelister" and "skattelistene," which both translate literally to "tax list." In this category, we also include searches for the words "skattelister" and "skattelistene," which both translate literally to "tax list." Also included in this category are searches for the phrases "skattelister" and "skattelistene," which literally translate to "tax list" and "tax list list," respectively. In addition, searches for the terms "skattelister" and "skattelistene" are included in this section, which are both translated literally as "tax list" or "tax list list." Websites such as www.skattelister.no, which provided access to tax information, were well-known in Norway at the time of their creation. For our study, we will use statistics from two terms that have regularly been among the most popular in the world: "weather" and "YouTube." These two phrases, which are both related to the weather, will serve as the basis for our research. It is feasible to determine how much the general public is interested in learning about taxes based on the number of searches for the word "tax" on Google.

A bar plot, like depicted in the previous paragraph, depicts the popularity of chosen phrases in Norway in 2010. In addition, this report includes a line graph illustrating the distribution of Google searches over the course of each week in the year 2010. If you're searching during the first week of 2010, search volumes are normalized such that the total number of searches across all categories equals one in total.

@@

Table 1 shows the result of regression model using the same OLS regression. Given the dependent variable happiness, we can see that the estimated coefficient on Income Rank is 0.120, which is positive. Moreover, the p-value of the coefficient is smaller than 0.05, which indicates that Income Rank as explanatory variable is statistical significant to our dependent variable (happiness). This coefficient implies that going from the lowest to the highest income rank in Norway during 1985–2000 was associated with an increase in happiness of 0.120 standard deviations. In addition, this table reports that the coefficient of the interaction term is 0.090. The p-value of 0.032 also implies that this variable is statistical significant to our dependent variable happiness.

@@

In this paper, we compare the popularity of Norwegian tax lists with YouTube searches from January 2010 to January 2011. This paper will also test whether the increase in income transparency affects the gap between richer and poorer people's well-being. Figure 1 shows that a skattelister in Norway was able to find tax documents on the Internet by inserting the word "YouTube". This indicates that a great deal of interest in the tax listings existed. Figure 2 shows that the first week of 2010 saw a rise in the overall number of searches across all categories, with the number of searches for weather-related information almost doubling. Table 1 shows that going from the lowest income rank in Norway to the highest income rank in Norway during 1985 - 2000 was associated with a 0.120 standard deviation increase in happiness. The online tax lists of Norwegian taxpayers reveal income including gains outside of the payroll, which makes the data useless for salary comparisons practices.

@@
